Get population and temperature data, version 2

### 2021-11-27 getdata2.Rmd
first commit

---
title: "Get population and temperature data"
author: "Göran Broström"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

@@

@@

```{r readobs}
bb <- obs$id[obs$starttyp == 2]
births <- obs[obs$id %in% bb, 
    c("id", "sex", "birthdate", "nofrs", "sluttyp", "socStatus", "hisclass", 
      "enter", "exit", "ort", "ortnmn")]
births$event <- as.numeric(births$sluttyp == 2)
births <- age.window(births, c(0, 1))
births <- births[order(births$id, births$enter), ]
##indx <- tapply(births$id, births$id)
##births$mexit <- tapply(births$exit, births$id, max)[indx]
##births$mevent <- tapply(births$event, births$id, max)[indx]
##births <- births[!duplicated(births$id), ]
### Deaths:
did <- unique(births$id[births$event == 1])
deaths <- births[births$id %in% did, ]
indx <- tapply(deaths$id, deaths$id)
deaths$menter <- tapply(deaths$enter, deaths$id, min)[indx]
deaths <- deaths[!duplicated(deaths$id), ]
deaths$enter <- deaths$menter
deaths$menter <- NULL
deaths$enter <- 0 ## satisfied....
### Survivors:
survs <- births[!(births$id %in% did), ]
indx <- tapply(survs$id, survs$id)
survs$mexit <- tapply(survs$exit, survs$id, max)[indx]
survs <- survs[!duplicated(survs$id), ]
survs$exit <- survs$mexit
survs$mexit <- NULL
survs$enter <- 0 ## Ok ...
### Now, give birth to NEW births:
births <- rbind(deaths, survs)
births <- births[births$birthdate > 1895, ]

@@

```{r neo, fig.cap = "Neonatal mortality by year, all regions."}
neo <- age.window(births, c(0, 1 / 13))
surv <- as.matrix(neo[, c("enter", "exit", "event", "birthdate")])
pall <- perstat(surv, period = 1900:1951)$intensity
plot(1900:1950, pall, type = "b", ylim = c(0, 0.5), col = "blue", xlab = "Year",
     ylab = "Mortality")
abline(h = 0)
abline(h = 0.3, lty = 3)
```

>>>>

---
title: "Get population and temperature data, version 2"
author: "Göran Broström"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

@@

# Introduction

The only (so far) difference compared to the article "getdata.Rmd" is that time 
here is measured in *days* instead of in a fraction of a year.

@@

```{r readobs}
bb <- obs$id[obs$starttyp == 2]
births <- obs[obs$id %in% bb, 
    c("id", "sex", "foddat", "startdat", "slutdat", "nofrs", "sluttyp", "socStatus", "hisclass", 
      "ort", "ortnmn")]
births$event <- as.numeric(births$sluttyp == 2)
## New start in days...:
births$enter <- as.numeric(with(births, difftime(startdat, foddat, unit = "days")))
births$exit <- as.numeric(with(births, difftime(slutdat, foddat, unit = "days")))
births <- age.window(births, c(0, 365))
births <- births[order(births$id, births$enter), ]
##indx <- tapply(births$id, births$id)
##births$mexit <- tapply(births$exit, births$id, max)[indx]
##births$mevent <- tapply(births$event, births$id, max)[indx]
##births <- births[!duplicated(births$id), ]
### Deaths:
did <- unique(births$id[births$event == 1])
deaths <- births[births$id %in% did, ]
indx <- tapply(deaths$id, deaths$id)
deaths$menter <- tapply(deaths$enter, deaths$id, min)[indx]
deaths <- deaths[!duplicated(deaths$id), ]
deaths$enter <- deaths$menter
deaths$menter <- NULL
deaths$enter <- 0 ## satisfied....
### Survivors:
survs <- births[!(births$id %in% did), ]
indx <- tapply(survs$id, survs$id)
survs$mexit <- tapply(survs$exit, survs$id, max)[indx]
survs <- survs[!duplicated(survs$id), ]
survs$exit <- survs$mexit
survs$mexit <- NULL
survs$enter <- 0 ## Ok ...
### Now, give birth to NEW births:
births <- rbind(deaths, survs)
births <- births[births$foddat > as.Date("1894-12-31"), ]

@@

```{r neo, fig.cap = "Neonatal mortality by year, all regions."}
neo <- births
neo$exit <- neo$exit / 365
neo$birthdate <- toTime(neo$foddat)
neo <- age.window(neo, c(0, 28 / 365))
##neo$exit <- 365 * neo$exit
surv <- as.matrix(neo[, c("enter", "exit", "event", "birthdate")])
pall <- perstat(surv, period = 1900:1951)$intensity
plot(1900:1950, pall, type = "b", ylim = c(0, 0.5), col = "blue", xlab = "Year",
     ylab = "Mortality")
abline(h = 0)
abline(h = 0.3, lty = 3)
```
